# Tour of AnVIL

## Introduction

AnVIL consists of several different pieces that work together to provide researchers with the things they need to carry out their analyses.

```{r, echo=FALSE, out.width="80%", fig.alt="Diagram showing different parts of the AnVIL ecosystem.", fig.align="center"}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1vJebvBQH_U80gDBvj_1d0USMnhaB3jiavWTLQDhgRuw/edit?slide=id.h66921c4e16512379_0_119#slide=id.h66921c4e16512379_0_119" )
```


It can be easy to start thinking of AnVIL as the just computing platform (Terra), and that is indeed where you will spend the majority of your time. But AnVIL also includes:

- The AnVIL Data Catalog, providing secure access to a large collection of high-interest genomic datasets, along with the [AnVIL Data Explorer](https://explore.anvilproject.org/datasets) to help you find and import these datasets into your own research.
- [Dockstore](https://dockstore.org/) for finding and sharing analysis pipelines, which can be imported into AnVIL with the click of a button.
- The [community support forum](https://help.anvilproject.org/), where you can interact with other AnVIL users as well as members of the AnVIL team to find help and hear about what people are doing on AnVIL.
- The [AnVIL Portal](https://anvilproject.org/), AnVIL’s primary website, containing documentation and FAQS, announcements about new features and datasets, upcoming events, and more.

In this chapter we will take a brief tour of the broader AnVIL ecosystem, showing you how to access these resources, before diving into the Terra interface.

## The AnVIL Ecosystem

### AnVIL Portal

The AnVIL Portal ([anvilproject.org](https://anvilproject.org/)) is the central hub of the AnVIL ecosystem. Here you can find links to the many components of AnVIL, as well as documentation and important announcements.

```{r, echo=FALSE, out.width="80%", fig.alt="Screenshot of the AnVIL Portal home page.", fig.align="center"}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1vJebvBQH_U80gDBvj_1d0USMnhaB3jiavWTLQDhgRuw/edit?slide=id.g3fb902435d4_1_4#slide=id.g3fb902435d4_1_4" )
```


We’ve highlighted a few items that may be useful to you:

- Click on the cards to access important AnVIL tools including Terra and Dockstore, which we will discuss in more detail below.
- Check the menu at the top of the page for:
  - Learn contains documentation and tutorials
  - AnVIL Data Explorer lets you find and access datasets hosted by AnVIL

```{r, echo=FALSE, out.width="80%", fig.alt="Screenshot of the 'Learn' page on the AnVIL Portal.", fig.align="center"}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1vJebvBQH_U80gDBvj_1d0USMnhaB3jiavWTLQDhgRuw/edit?slide=id.g3fb902435d4_1_20#slide=id.g3fb902435d4_1_20" )
```

At [anvilproject.org/learn](http://anvilproject.org/learn), you can find documentation and tutorials for many of AnVIL’s features.

:::{.reflection}
TODO
:::

